Peaks Over Threshold for Bursty Time Series

### 2018-04-23 article_springer/article_springer.Rmd
Change CTRM => CTRE and ctrm => ctre throughout

---
title: Peaks Over Threshold for Bursty Time Series
# titlerunning: Inference for Heavy-Tailed CTRMs
thanks: |
    Peter Straka was supported by the 
    Discovery Early Career Research Award DE160101147 on the Project 
    "Predicting Extremes when Events Occur in Bursts" by the Australian
    Research Council.
    Katharina Hees was supported by 
    the DAAD co-financed by the German Federal Ministry of Education and 
    Research (BMBF).

@@

```{r get-rticles, eval=FALSE, include=FALSE, cache=TRUE}
# use this rticles version: 
devtools::install_github("strakaps/rticles", 
                         ref = "724aeb88d005e2f7befc37328633e7cb2bc1cce0")
# use the newest MittagLeffleR version:
devtools::install_github("strakaps/MittagLeffleR")
library(MittagLeffleR)
devtools::install_github("UNSW-MATH/CTRM")
library(CTRM)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  cache = TRUE,
  message = FALSE,
  fig.height = 3, out.width = '\\textwidth'
  )
library(CTRM)
library(MittagLeffleR)
library(magrittr)
```

@@

We review the marked renewal process in Section 2, and 
derive a scaling limit theorem for inter-exceedance times in Section 3.
We give a statistical procedure to estimate model parameters via stability plots 
in Section 5, but to set the stage we first discuss inference for the 
Mittag-Leffler distribution in Section 4. 
Diagnostic plots for model criticism are discussed in Section 6. 
In Section 7, we discuss the non-memorylessness property of the 
Mittag-Leffler distribution, and how it affects the predictive distribution 
for the next threshold crossing event. 
Section 8 concludes.
For all statistical computations we have used R [@R]. 
Code used for the analysis in this article has been organized into 
an R package `CTRM` (<https://github.com/UNSW-MATH/CTRM>).
The source code for the figures generated in this manuscript is 
available online at <https://github.com/UNSW-MATH/bursty-POT>.

@@

```{r thresholdedBursty, message=FALSE, fig.height=7, fig.width=9, fig.cap="\\label{fig:thresholdedBursty}Exceedances (red) and Times until Exceedance (durations between blue crosses) for a given threshold $\\ell$ (dashed line)."}
tail <- params$tail
n <- params$n 
sigma <- (cos(pi*tail/2))^(1/tail)
times <- cumsum(stabledist::rstable(n, tail, 1, sigma, pm=1))
#times<-cumsum(rpareto(n,tail))
magnitudes <- extRemes::revd(n, scale = 1, shape = 0)
sim_ctrm <- ctrm(data.frame(times, magnitudes)) 
par(mfrow = c(2,1))
plot(sim_ctrm, p = 0.01, main = "Simulated MRP") 
flares %>% ctrm() %>% plot(p = 0.02, log = 'y', main = "HXRBS data") 
```

@@

flare_mags <- flares %>% ctrm() %>% thin(k=500) %>% magnitudes()
flares_exponent <- flare_mags %>% qqestplot_static()

@@

```{r likelihood-ratio-test}
flares_arrivals <- flares %>% ctrm() %>% thin(k = 500) %>% interarrival()
mlmle_out <- mlmle(data = flares_arrivals)
```

@@

```{r solar-flare-tail-scale, message=FALSE, fig.height=4.5, fig.width=8, fig.cap="\\label{fig:flares}Stability plots for the tail and scale parameter of the Mittag-Leffler distribution of the Solar Flare dataset. Dotted horizontal lines are at $\\beta = 0.85$ and $\\sigma_0 = 3 \\times 10^7$ seconds $\\approx 0.95$ years."}
thin_flares_ctrm <- flares %>% ctrm() %>% thin(k = 700)
par(mfrow=c(1,2))
MLestimates(thin_flares_ctrm, tail = 0.9, scale = 3 * 1E7)
```

@@

```{r flare-diagnostics-1, fig.height=7, fig.cap="Diagnostic plots for the solar flare data: auto-correlation function.\\label{fig:flare-diagnostics-1}"}
flares_ctrm <- flares %>% ctrm() %>% thin(k = 150)
acf(flares_ctrm)
```

```{r flare-diagnostics-2, fig.height=7, fig.cap="Diagnostic plots for the solar flare data: empirical copula and QQ Plot. \\label{fig:flare-diagnostics-2}"}
par(mfrow = c(2,2))
empcopula(flares_ctrm)
flares_ctrm %>% interarrival() %>% mlqqplot(log = 'xy', tail = 0.8, main = "Mittag-Leffler QQ Plot")
```

@@

```{r simulated-example, message=FALSE, fig.cap=caption}
k_max <- 500 
thin_sim_ctrm <- thin(ctrm = sim_ctrm, k = k_max)
par(mfrow=c(1,2))
MLestimates(ctrm = thin_sim_ctrm, tail = params$tail, scale = n^(1/params$tail))
```

>>>>

---
title: "Peaks Over Threshold for Bursty Time Series"
# titlerunning: "CTRE: POT for Bursty Time Series"
thanks: |
    Peter Straka was supported by the 
    Discovery Early Career Research Award DE160101147 on the Project 
    "Predicting Extremes when Events Occur in Bursts" by the Australian
    Research Council.
    Katharina Hees was supported by 
    the DAAD co-financed by the German Federal Ministry of Education and 
    Research (BMBF).

@@

```{r get-rticles, eval=FALSE, include=FALSE, cache=TRUE}
# use this rticles version: 
devtools::install_github("strakaps/rticles", 
                         ref = "724aeb88d005e2f7befc37328633e7cb2bc1cce0")
# use the newest MittagLeffleR version:
devtools::install_github("strakaps/MittagLeffleR")
library(MittagLeffleR)
devtools::install_github("UNSW-MATH/CTRE")
library(CTRE)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  cache = TRUE,
  message = FALSE,
  fig.height = 3, out.width = '\\textwidth'
  )
library(CTRE)
library(MittagLeffleR)
library(magrittr)
```

@@

We review the marked renewal process in Section 2, and 
derive a scaling limit theorem for inter-exceedance times in Section 3.
We give a statistical procedure to estimate model parameters via stability plots 
in Section 5, but to set the stage we first discuss inference for the 
Mittag-Leffler distribution in Section 4. 
Diagnostic plots for model criticism are discussed in Section 6. 
In Section 7, we discuss the non-memorylessness property of the 
Mittag-Leffler distribution, and how it affects the predictive distribution 
for the next threshold crossing event. 
Section 8 concludes.
For all statistical computations we have used R [@R]. 
Code used for the analysis in this article has been organized into 
an R package `CTRE` (<https://github.com/UNSW-MATH/CTRE>).
The source code for the figures generated in this manuscript is 
available online at <https://github.com/UNSW-MATH/bursty-POT>.

@@

```{r thresholdedBursty, message=FALSE, fig.height=7, fig.width=9, fig.cap="\\label{fig:thresholdedBursty}Exceedances (red) and Times until Exceedance (durations between blue crosses) for a given threshold $\\ell$ (dashed line)."}
tail <- params$tail
n <- params$n 
sigma <- (cos(pi*tail/2))^(1/tail)
times <- cumsum(stabledist::rstable(n, tail, 1, sigma, pm=1))
#times<-cumsum(rpareto(n,tail))
magnitudes <- extRemes::revd(n, scale = 1, shape = 0)
sim_ctre <- ctre(data.frame(times, magnitudes)) 
par(mfrow = c(2,1))
plot(sim_ctre, p = 0.01, main = "Simulated MRP") 
flares %>% ctre() %>% plot(p = 0.02, log = 'y', main = "HXRBS data") 
```

@@

flare_mags <- flares %>% ctre() %>% thin(k=500) %>% magnitudes()
flares_exponent <- flare_mags %>% qqestplot_static()

@@

```{r likelihood-ratio-test}
flares_arrivals <- flares %>% ctre() %>% thin(k = 500) %>% interarrival()
mlmle_out <- mlmle(data = flares_arrivals)
```

@@

```{r solar-flare-tail-scale, message=FALSE, fig.height=4.5, fig.width=8, fig.cap="\\label{fig:flares}Stability plots for the tail and scale parameter of the Mittag-Leffler distribution of the Solar Flare dataset. Dotted horizontal lines are at $\\beta = 0.85$ and $\\sigma_0 = 3 \\times 10^7$ seconds $\\approx 0.95$ years."}
thin_flares_ctre <- flares %>% ctre() %>% thin(k = 700)
par(mfrow=c(1,2))
MLestimates(thin_flares_ctre, tail = 0.9, scale = 3 * 1E7)
```

@@

```{r flare-diagnostics-1, fig.height=7, fig.cap="Diagnostic plots for the solar flare data: auto-correlation function.\\label{fig:flare-diagnostics-1}"}
flares_ctre <- flares %>% ctre() %>% thin(k = 150)
acf(flares_ctre)
```

```{r flare-diagnostics-2, fig.height=7, fig.cap="Diagnostic plots for the solar flare data: empirical copula and QQ Plot. \\label{fig:flare-diagnostics-2}"}
par(mfrow = c(2,2))
empcopula(flares_ctre)
flares_ctre %>% interarrival() %>% mlqqplot(log = 'xy', tail = 0.8, main = "Mittag-Leffler QQ Plot")
```

@@

```{r simulated-example, message=FALSE, fig.cap=caption}
k_max <- 500 
thin_sim_ctre <- thin(ctre = sim_ctre, k = k_max)
par(mfrow=c(1,2))
MLestimates(ctre = thin_sim_ctre, tail = params$tail, scale = n^(1/params$tail))
```
